Statement-level AST-based Clone Detection in Java using Resolved Symbols

### 2019-11-07 benevol/paper.tex
Updates from Overleaf

We propose a method to efficiently detect clones taking into account contextual information. We introduce a tool that uses an AST-parsing library named JavaParser to detect clones and retrieve contextual information. Our method parses the Abstract Syntax Tree retrieved from JavaParser into a graph structure, which is used to find clones. This graph maps the following relations for each statement in the codebase: the next statement, the previous statement, and the previous cloned statement.
\end{abstract}

@@

\section{Introduction}
%Most clone detection tools can be configured using thresholds. These thresholds indicate the minimum number of lines, tokens and/or statements that must be spanned for duplicate fragments to be considered clones. Often, such thresholds are intuitively chosen~\cite{li2006cp, roy2009mutation} or based on a quartile distribution of empirical data~\cite{alves2010deriving}. Using the maintainability score we can find support for which thresholds should be chosen to increase the chance to find clones that improve maintainability when refactored.
Duplicate code fragments are often considered as symptoms of bad design~\cite{fowler2018refactoring}. They create implicit dependencies, thus increasing maintenance efforts or causing bugs in evolving software. Changing one occurrence of a duplicated fragment may require changes in other occurrences ~\cite{ostberg2014automatically}. Also, duplicated code was shown to add up to 25\% of total system volume~\cite{bruntink2005use}, which entails more code to be maintained.

A lot of tools have been proposed to detect such duplication issues \cite{roy2009comparison, svajlenko2014evaluating, sheneamer2016survey}. These tools can find matching fragments of code, but do not take into account contextual information of code. An example of such contextual information is the name of used variables: many different methods with the same name can exist in a codebase. This can obstruct refactoring opportunities.

We describe a method to detect clones while taking into account such contextual information and propose a tool to detect clones taking into account this contextual information. Run this tool over a corpus of 2,267 Java projects. We find that taking into account contextual information results in 11\% less clones than found when not taking this information into account. Manually inspecting a sample the difference, we find that all clones with different contextual information in our sample are less relevant for refactoring.

\section{Background}
\subsection{Code clones}
We use two concepts to argue about code clones~\cite{roy2007survey}:
\\ \textbf{Clone instance}: A single cloned fragment.
\\ \textbf{Clone class}: A set of similar clone instances.

\section{JavaParser}
JavaParser~\cite{tomassetti2017javaparser} is a Java library which allows parsing Java source files to an abstract syntax tree (AST). Integrated into JavaParser is a library named SymbolSolver. This library allows for the resolution of symbols using JavaParser. For instance, we can use it to trace references (methods, variables, types, etc) to their declarations (these referenced identifiers are also called ``symbols''). Using this, we can find the required contextual information.

\section{Motivating example} \label{sec:type1}
\begin{figure*}
\begin{parcolumns}{2}
\colchunk[1]{
\begin{javacode}
package com.sb.game;

@@

Most clone detection tools~\cite{kamiya2002ccfinder, semura2017ccfindersw, roy2008nicad, svajlenko2016bigcloneeval, svajlenko2014evaluating} detect type 1 clones by textually comparing code fragments (except for whitespace and comments). Although textually equal, method, type and variable references can still refer to different declarations. In such cases, refactoring opportunities could be invalidated. This can make clones less suitable for refactoring purposes, as they require additional judgment regarding the refactorability of such a clone.

Figure~\ref{fig:type1} shows two clone classes. Merging these clone classes is very hard (and likely not desirable), as both cloned fragments describe different functional behavior. The first cloned fragment is a method that adds something to a \texttt{List}. However, the \texttt{List} objects to which something is added are different. Looking at the \texttt{import} statement above the class, one fragment uses the \texttt{java.util.List} and the other uses the \texttt{java.awt.List}. Both happen to have an \texttt{add} method, but apart from that their implementation is completely different.

@@

This shows that not all textually equal clones can easily be refactored.

@@

We compare the fully qualified method signature for all method references. A fully qualified method signature consists of the fully qualified name of the method, the fully qualified type of the method plus the fully qualified type of each of its arguments. For instance, an \texttt{eat} method could become \texttt{com.sb.AppleCore com.sb.fruitgame.Apple.eat(com.sb.fruitgame.Tool)}.

@@

\section{Results}
To compare the difference in detected clones when contextual information is considered, we compared the number of clones found when considering contextual information with when it is not considered. For this, we use a corpus of 2.

\section{Conclusion}

>>>>

We propose a method to efficiently detect clones taking into account contextual information. We introduce a tool that uses an AST-parsing library named JavaParser to detect clones and retrieve contextual information. Our method parses the Abstract Syntax Tree retrieved from JavaParser into a graph structure, which is used to find clones. This graph maps the following relations for each statement in the codebase: the next statement, the previous statement, and the previous cloned statement.

We find that, when taking into account contextual information in our clone detection, we find 11\% fewer clones. Manually inspecting the difference, we find that they are less relevant for refactoring.
\end{abstract}

@@

\section{Introduction}
%Most clone detection tools can be configured using thresholds. These thresholds indicate the minimum number of lines, tokens and/or statements that must be spanned for duplicate fragments to be considered clones. Often, such thresholds are intuitively chosen~\cite{li2006cp, roy2009mutation} or based on a quartile distribution of empirical data~\cite{alves2010deriving}. Using the maintainability score we can find support for which thresholds should be chosen to increase the chance to find clones that improve maintainability when refactored.
Duplicate code fragments are often considered as symptoms of bad design~\cite{fowler2018refactoring}. They create implicit dependencies, thus increasing maintenance efforts or causing bugs in evolving software. Changing one occurrence of a duplicated fragment may require changes in other occurrences ~\cite{ostberg2014automatically}. Also, duplicated code was shown to account for 25\% of total system volume~\cite{bruntink2005use}, entailing more code to be maintained.

Several tools have been proposed to detect duplication issues \cite{roy2009comparison, svajlenko2014evaluating, sheneamer2016survey}. These tools can find matching fragments of code, however do not take into account contextual information of code. An example of such contextual information is the name of used variables: many different methods with the same name can exist in a codebase. This can obstruct refactoring opportunities.

We describe a method to detect clones while taking into account the contextual information and introduce a tool to detect clones taking into account this contextual information. We run the tool over a corpus of 2,267 Java projects. We find that taking into account contextual information results in 11\% less clones than found when not taking this information into account. Manually inspecting a sample the difference, we find that all clones with different contextual information in our sample are less relevant for refactoring.

\begin{figure*}
\begin{parcolumns}{2}
\colchunk[1]{
\begin{javacode}
package com.sb.game;

@@

\section{Background}
We first explain relevant code clone terminology and definitions. Next, we describe the JavaParser tool, used to retrieve contextual information of source code.

\subsection{Code clones}
We use two concepts to argue about code clones~\cite{roy2007survey}:
\\ \textbf{Clone instance}: A single cloned fragment.
\\ \textbf{Clone class}: A set of similar clone instances.

\subsection{JavaParser}
JavaParser~\cite{tomassetti2017javaparser} is a Java library which allows parsing Java source files to an abstract syntax tree (AST). Integrated into JavaParser is a library named SymbolSolver. This library allows for the resolution of symbols using JavaParser. For instance, we can use it to trace references (methods, variables, types, etc) to their declarations (these referenced identifiers are also called ``symbols''). Using this, we can find the required contextual information.

\section{Motivating example} \label{sec:type1}
Most clone detection tools~\cite{kamiya2002ccfinder, semura2017ccfindersw, roy2008nicad, svajlenko2016bigcloneeval, svajlenko2014evaluating} detect type 1 clones by textually comparing code fragments (except for whitespace and comments). Although textually equal, method, type and variable references can still refer to different declarations. In such cases, refactoring opportunities could be invalidated. This can make clones less suitable for refactoring purposes, as they require additional judgment regarding the refactorability of such a clone.

Figure~\ref{fig:type1} shows two clone classes. Merging these clone classes is very hard (and likely not desirable), as both cloned fragments describe different functional behavior. The first cloned fragment is a method that adds something to a \texttt{List}. However, the \texttt{List} objects to which something is added are different. Looking at the \texttt{import} statement above the class, one fragment uses the \texttt{java.util.List} and the other uses the \texttt{java.awt.List}. Both happen to have an \texttt{add} method, however their implementation is completely different.

@@

This shows that not all textually equal clones can be easily refactored.

@@

We compare the fully qualified method signature for all method references. A fully qualified method signature consists of the fully qualified name of the method, the fully qualified type of the method plus the fully qualified type of each of its arguments. For instance, an \texttt{eat} method could become 
%\footnotesize{
\texttt{com.sb.Apple.eat(com.sb.Tool)}.
%}.

@@

\section{Experiments and Results}
To compare the difference in detected clones when contextual information is considered, we compared the number of clones found when considering contextual information with when it is not considered. For this, we use a corpus of 2,267 Java projects including their dependencies \cite{baars2019towards}.

We find that 167,913 clones are found when contextual information is not considered, whereas 149,569 clones are found when it is considered. We manually analyse a sample of 50 clones that are not found when considering contextual information. We found that these clones are indeed hard to refactor because they describe different functional operations and can thus not be extracted to a new method. Also, most often they were not relevant because, based on our judgement, refactoring would not improve the code design. Often, different methods were called or variables of different types were used.

We also look into the difference in performance when taking into account contextual information. Detecting the clones in all 2,267 projects took 20.83 minutes when considering contextual information. When contextual information was not considered, it took 1.58 minutes.

\section{Conclusion}
We propose a method to detect clones taking into account contextual information of source code. Contextual information is important because different functional concepts may turn up as clones because they are textually identical.

We define three aspects of source code as the contextual information:
\begin{itemize}
    \item The type of variable references.
    \item The location of method references.
    \item The location of type references.
\end{itemize}
When these references have the same name but point to different locations, clones may not be easily refactorable. Our results show that most such clones are not relevant for refactoring. This accounts for about 11\% of clones. This comes however with a performance trade-off: detecting clones with contextual information took 13 times longer than when not taking contextual information into account (1.6min vs 20.8min).
